Sequential rank agreement methods for comparison of ranked lists

### 2015-08-18 latex/paper.Rnw
Code fix
Adjusting figure margins
Fixing figure margins
homogenizing graphics

\title{Sequential rank agreement methods for comparison of lists}
\author{Claus Thorn Ekstrøm}
\author{Thomas Alexander Gerds}
\author{Andreas Kryger Jensen}
\author{Kasper Brink-Jensen}
\affil{Biostatistics, University of Copenhagen}
\renewcommand{\topfraction}{.85}
\renewcommand{\bottomfraction}{.7}
\renewcommand{\textfraction}{.15}
\renewcommand{\floatpagefraction}{.66}
\renewcommand{\dbltopfraction}{.66}
\renewcommand{\dblfloatpagefraction}{.66}
\setcounter{topnumber}{9}
\setcounter{bottomnumber}{9}
\setcounter{totalnumber}{20}
\setcounter{dbltopnumber}{9}

@@

options(SweaveHooks=list(fig=function() par(cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5)))

@@

<<fig1a, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE,eval=TRUE>>=
ysize <- 1000
nreference <- 20
## Plot the line
plot(res[1:ysize], col="black", lwd=3, type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")

@@

<<fig1b, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE, eval=TRUE>>=
## Same as above but zoomed in
ysize <- 50
plot(res[1:ysize], lwd=3, col="black", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")

@@

<<fig2a, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE, eval=TRUE>>=
ysize <- 1000
nreference <- 20
## Plot the line
plot(res2[1:ysize], col="black", lwd=3, type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")

@@

## Load the data for the methods
makereference <- function(x, y, n=100, B=20) {
    sapply(1:n, function(i) { mymat <- producelists(x, sample(y)) ; mymat[(topk+1):nrow(mymat),] <- NA ; sra(mymat, B=B) }  )
}
## Uncomment the two lines below to redu the analysis
##null2 <- makereference(x, y, n=100, B=B)
##save(null2, file="R/fig1null2c.rda")
load("R/fig1null2c.rda")

@@

<<fig2b, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE, eval=TRUE>>=
## Same as above but zoomed in
ysize <- 50
plot(res2[1:ysize], lwd=3, col="black", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")

@@

\begin{figure}[tb]
   \begin{center}
 \includegraphics[width=.45\textwidth]{paper-fig1a}
 \includegraphics[width=.45\textwidth]{paper-fig2a}
 \includegraphics[width=.45\textwidth]{paper-fig1b}
 \includegraphics[width=.45\textwidth]{paper-fig2b}
 \end{center}
 \caption{Left panels: Sequential rank agreement for 4 different
   analysis methods applied to the 3051 genes in the Golub data /black
   line). Right
   panels: Corresponding sequential rank agreement for the same data
   but where only the top 20 ranked items are available. The blue and
   red areas correspond to the independent and randomized reference
   hypothesis areas, respectively. The bottom plots are identical to
   the corresponding top plots but have been zoomed in on the first
   part of the $x$ axis.}
 \label{fig:example1}
 \end{figure}

@@

<<fig-ao1, echo=FALSE, fig=TRUE,include=FALSE, eval=TRUE>>=
myres <- average_overlap(inputmatrix[,1:2])
myres2 <- average_overlap(inputmatrix[-c(1,2),1:2])
xvals <- 1:500
plot(xvals, myres[xvals], xlim=c(1, 150), ylim=c(0,1), type="l", lwd=2, ylab="Average overlap", xlab="List depth")
lines(xvals[-c(1,2)], myres2[1:498], col="red", lwd=2)
@

<<fig-ao2, echo=FALSE, fig=TRUE,include=FALSE, eval=TRUE>>=
mmyres <- sra(inputmatrix[,1:2])
mmyres2 <- sra(inputmatrix[-c(1,2),1:2])
xvals <- 1:500
plot(xvals, mmyres[xvals], xlim=c(1, 150), ylim=c(0, 110), type="l", lwd=2, ylab="Sequential rank agreement", xlab="List depth")
lines(xvals[-c(1,2)], mmyres2[1:498], col="red", lwd=2)
@

\begin{figure}[tb]
\begin{center}
\includegraphics[width=.45\textwidth]{paper-fig-ao1}
\includegraphics[width=.45\textwidth]{paper-fig-ao2}
\end{center}
 \caption{Average overlap (left figure) and sequential rank agreement
 (right figure) for comparing two different analysis methods
 (marginal $t$ test and logistic regression)
    applied to the Golub data. The black lines are based on the full
    set of predictors while the red lines have the two highest
    associated predictors removed
    from the data before computing the average overlap and sequential
    rank agreement.}
  \label{fig:case1}
 \end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=.5\textwidth]{pics/riskAgreementPlot}%
\includegraphics[width=.5\textwidth]{pics/predictorAgreementPlot}
\end{center}
\caption{Left panel: Sequential rank agreement of 1000 rankings of the
  predicted risks of malignant tumor. For each method the different
  rankings were obtained by first training models in 1000 random
  sub-samples of the MALOVA data and then predicting the risk of
  malignant tumor in the 113 DACOVA patients. Right panel: Sequential
  rank agreement of 1000 rankings of the 5000 predictors. The rankings
  were obtained from the same 1000 trained models.}
 \label{fig:app1}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=.5\textwidth]{pics/ridgeCensoring}%
\includegraphics[width=.5\textwidth]{pics/aucPlot}
\end{center}
\caption{Left panel: Sequential rank agreement for Ridge regression
  obtained by artificially censoring predictors when their absolute
  coefficient values are lower than the 0.1\% quantile. Right panel:
  Box plots of AUC values across the $1000$ sub-samples with respect
  to the known class labels of the DACOVA data.}
 \label{fig:app2}
\end{figure}

>>>>

\title{Sequential rank agreement methods for comparison of ranked lists}
\author{Claus Thorn Ekstrøm}
\author{Thomas Alexander Gerds}
\author{Andreas Kryger Jensen}
\author{Kasper Brink-Jensen}
\affil{Biostatistics, University of Copenhagen}
\renewcommand{\topfraction}{.85}
\renewcommand{\bottomfraction}{.7}
\renewcommand{\textfraction}{.15}
\renewcommand{\floatpagefraction}{.66}
\renewcommand{\dbltopfraction}{.66}
\renewcommand{\dblfloatpagefraction}{.66}
\setcounter{topnumber}{9}
\setcounter{bottomnumber}{9}
\setcounter{totalnumber}{20}
\setcounter{dbltopnumber}{9}

@@

options(SweaveHooks=list(fig=function() par(cex.lab=1.4, cex.axis=1.5, cex.main=1.2, cex.sub=1.5)))

@@

<<fig1a, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE,eval=TRUE>>=
ysize <- 1000
nreference <- 20
## Plot the line
plot(res[1:ysize], col="black", lwd=3, type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="List depth", bty="n")

@@

<<fig1b, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE, eval=TRUE>>=
## Same as above but zoomed in
ysize <- 50
plot(res[1:ysize], lwd=3, col="black", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="List depth", bty="n")

@@

<<fig2a, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE, eval=TRUE>>=
ysize <- 1000
nreference <- 20
## Plot the line
plot(res2[1:ysize], col="black", lwd=3, type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="List depth", bty="n")

@@

## Load the data for the methods
makereference <- function(x, y, n=100, B=20) {
    sapply(1:n, function(i) { mymat <- producelists(x, sample(y)) ; mymat[(topk+1):nrow(mymat),] <- NA ; sra(mymat, B=B) }  )
}
## Uncomment the two lines below to redo the analysis
##null2 <- makereference(x, y, n=100, B=B)
##save(null2, file="R/fig1null2c.rda")
load("R/fig1null2c.rda")

@@

<<fig2b, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE, eval=TRUE>>=
## Same as above but zoomed in
ysize <- 50
plot(res2[1:ysize], lwd=3, col="black", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="List depth", bty="n")

@@

\begin{figure}[htbp]
   \begin{center}
 \includegraphics[width=.49\textwidth]{paper-fig1a}
 \includegraphics[width=.49\textwidth]{paper-fig2a}
 \includegraphics[width=.49\textwidth]{paper-fig1b}
 \includegraphics[width=.49\textwidth]{paper-fig2b}
 \end{center}
 \caption{Left panels: Sequential rank agreement for 4 different
   analysis methods applied to the 3051 genes in the Golub data /black
   line). Right
   panels: Corresponding sequential rank agreement for the same data
   but where only the top 20 ranked items are available. The blue and
   red areas correspond to the independent and randomized reference
   hypothesis areas, respectively. The bottom plots are identical to
   the corresponding top plots but have been zoomed in on the first
   part of the $x$ axis.}
 \label{fig:example1}
 \end{figure}

@@

<<fig-ao1, echo=FALSE, fig=TRUE,include=FALSE, eval=TRUE>>=
myres <- average_overlap(inputmatrix[,1:2])
myres2 <- average_overlap(inputmatrix[-c(1,2),1:2])
xvals <- 1:500
plot(xvals, myres[xvals], xlim=c(1, 150), ylim=c(0,1), type="l", lwd=2, ylab="Average overlap", xlab="List depth", bty="n")
lines(xvals[-c(1,2)], myres2[1:498], col="red", lwd=2)
@

<<fig-ao2, echo=FALSE, fig=TRUE,include=FALSE, eval=TRUE>>=
mmyres <- sra(inputmatrix[,1:2])
mmyres2 <- sra(inputmatrix[-c(1,2),1:2])
xvals <- 1:500
plot(xvals, mmyres[xvals], xlim=c(1, 150), ylim=c(0, 110), type="l", lwd=2, ylab="Sequential rank agreement", xlab="List depth", bty="n")
lines(xvals[-c(1,2)], mmyres2[1:498], col="red", lwd=2)
@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=.49\textwidth]{paper-fig-ao1}
\includegraphics[width=.49\textwidth]{paper-fig-ao2}
\end{center}
 \caption{Average overlap (left figure) and sequential rank agreement
 (right figure) for comparing two different analysis methods
 (marginal $t$ test and logistic regression)
    applied to the Golub data. The black lines are based on the full
    set of predictors while the red lines have the two highest
    associated predictors removed
    from the data before computing the average overlap and sequential
    rank agreement.}
  \label{fig:case1}
 \end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=.49\textwidth]{pics/riskAgreementPlot}%
\includegraphics[width=.49\textwidth]{pics/predictorAgreementPlot}
\end{center}
\caption{Left panel: Sequential rank agreement of 1000 rankings of the
  predicted risks of malignant tumor. For each method the different
  rankings were obtained by first training models in 1000 random
  sub-samples of the MALOVA data and then predicting the risk of
  malignant tumor in the 113 DACOVA patients. Right panel: Sequential
  rank agreement of 1000 rankings of the 5000 predictors. The rankings
  were obtained from the same 1000 trained models.}
 \label{fig:app1}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=.49\textwidth]{pics/ridgeCensoring}%
\includegraphics[width=.49\textwidth]{pics/aucPlot}
\end{center}
\caption{Left panel: Sequential rank agreement for Ridge regression
  obtained by artificially censoring predictors when their absolute
  coefficient values are lower than the 0.1\% quantile. Right panel:
  Box plots of AUC values across the $1000$ sub-samples with respect
  to the known class labels of the DACOVA data.}
 \label{fig:app2}
\end{figure}
